---
title: "`r params$title`"
author: "`r params$author`"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
output: html_document
params:
  title: NA
  author: NA
  fileraw: NA
  fileprep: NA 
  inputid: NA
  inputtarget: NA
  inputstartmet: NA
  colplot: NA
  inputnacol: NA
  inputnarow: NA
  inputimputmethod: NA
  filemetabolites: NA
  modeltraining: NA
  modeltesting: NA
  model_nfolds: NA
  model_nrepeats: NA
  model_ltune: NA
  modelcv: NA
  coeffs: NA
  resultscv: NA
  scoresall: NA

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(shiny)
library(rmarkdown)
library(knitr)
```

After loading the data, boxplot of the first columns of metabolites shows how the distributions is on the raw data.

```{r raw data}
cat("Raw data frame dimensions:", dim(params$fileraw))

raw_data_boxplots <- params$fileraw[,c(params$colplot)] %>%
  tidyr::gather(key = "metabolites", value = "value")#, starts_with(input$key_plot))

boxplot <- raw_data_boxplots %>%
  ggplot2::ggplot(aes(x = metabolites, y = value)) +
  ggplot2::geom_boxplot() +
  labs(x = "Metabolites", y = "Intensity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

boxplot
```

After preprocessing boxplot of the first columns of metabolites shows how the distributions after preprocessing.

```{r preprocessing data}
cat("Preprocessed data frame dimensions:", dim(params$fileprep))
cat("ID column:", params$inputid)
cat("Target column:", params$inputtarget)
cat("Threshold for na% column removal:", params$inputnacol)
cat("Threshold for na% row removal:", params$inputnarow)
cat("Imputation method:", params$inputimputmethod)


raw_data_boxplots <- params$fileprep[,c(params$colplot)] %>%
  tidyr::gather(key = "metabolites", value = "value")#, starts_with(input$key_plot))

boxplot <- raw_data_boxplots %>%
  ggplot2::ggplot(aes(x = metabolites, y = value)) +
  ggplot2::geom_boxplot() +
  labs(x = "Metabolites", y = "Intensity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

boxplot
```

For feature selection.

Single metabolite models. Meti \~ β0P + β1C + ... + Ԑ

The top metabolites are plotted as follow

```{r single metabolites}
params$filemetabolites

if(nrow(params$filemetabolites) >= 10){
  top_metabolites <- params$filemetabolites[1:10,]
  } else {
    top_metabolites <- params$filemetabolites
    }

met_pvalue <- top_metabolites %>%
  ggplot2::ggplot(aes(x = yvar, y = p.value_corrected)) +
  geom_segment( aes(x=yvar, xend=yvar, y=0, yend=p.value_corrected), color="black") +
  geom_point( color="gray", size=2, alpha=0.6) +
  theme_light() +
  coord_flip() +
  labs(x = "Metabolites", y = "p-value") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(axis.text.y = element_text(size = 5))

met_pvalue
```

ML testing

```{r ML}
params$modeltraining

params$modeltesting[-1]
```

ML with crossvalitation

```{r Multi-metabolite signature}
cat("Number of folds:", params$model_nfolds)
cat("Number of repeats:", params$model_nrepeats)
cat("tuneLenght used:", params$model_ltune)

params$modelcv

cat("Coefficients table \n")
params$coeffs

params$resultscv
```

Multi-metabolite scores

```{r ML score}
head(params$scoresall)

den <- params$scoresall %>%
            ggplot2::ggplot(aes(x = total)) +
            ggplot2::geom_density() +
            labs(x = "Total score", y = "density") +
            theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
          
den
```